Home

### 2025-07-20 about.qmd
Refactor website navigation and update content
Added a new 'People' link to the navigation in _quarto.yml and streamlined the 'Papers' section to a single link. Updated the 'About' page with a comprehensive overview of Professor Bin Yu's background and research focus, while enhancing the teaching section with detailed course descriptions for STAT 154 and STAT 215A. Adjusted styles.css to comment out the background color for the column class, improving layout consistency.

# About Bin

<img src="images/bio1.jpg" alt="Bin Yu speaking" style="float:right; margin:0 0 1em 2em; width:200px; border-radius:8px;" />

Bin Yu is CDSS Chancellor's Distinguished Professor in the Departments of statistics and EECS, and Center for Computational Biology, and serves as a scientific advisor at the Simons Institute for the Theory of Computing, all at UC Berkeley.

She obtained her BS Degree in Mathematics from Peking University, and MS and PhD Degrees in Statistics from UC Berkeley. She was Assistant Professor at UW-Madison, Visiting Assistant Professor at Yale University, Member of Technical Staff at Lucent Bell-Labs, and Miller Research Professor at Berkeley. She was a Visiting Faculty at MIT, ETH, Poincare Institute, Peking University, INRIA-Paris, Fields Institute at University of Toronto, Newton Institute at Cambridge University, and the Flatiron Institute in NYC. She was Chair of the Department of Statistics at UC Berkeley from 2009 to 2012. Recently, she was a 50% consultant researcher in the deep learning group of MSR at Redmond (2022-2023).

She is a Member of the U.S. National Academy of Sciences and of the American Academy of Arts and Sciences. She was President of the Institute of Mathematical Statistics (IMS), Guggenheim Fellow, Tukey Memorial Lecturer of the Bernoulli Society, Rietz Lecturer of IMS, and a COPSS E. L. Scott Prize winner. Recently, she delivered the Wald Memorial Lectures of IMS and COPSS Distinguished Achievement Award and Lecture (DAAL) (formerly Fisher Award and Lecture) at JSM in 2023. She served on the Inaugural Scientific Committee of the UK Turing Institute for Data Science and AI. She holds an Honorary Doctorate from the University of Lausanne in Switzerland. She is serving on the Editorial Board of Proceedings of National Academy of Sciences (PNAS).

@@

# About the Research Group

@@

@@

In 2014, I was elected to the National Academy of Sciences based on my statistical and scientific contributions, as well as my broad vision of data science best described in my article [Veridical Data Science](https://www.pnas.org/content/117/8/3920), written together with my former student Karl Kumbier. In this work, I introduced a framework based on three principles: predictability, computability and stability (abbreviated to PCS). This framework helps guide practitioners who solve domain data problems with data science tools to be creative in their analysis and properly validate their findings. I have written a book on the Veridical Data Science framework together with my former student [Rebecca Barter](https://www.rebeccabarter.com/). The book is being published by the MIT Press in 2024 with a free on-line copy available soon.

>>>>

# About Bin Yu

Bin Yu is the Chancellor's Distinguished Professor in the UC Berkeley Departments of statistics and EECS. She was Chair of the Department of Statistics at UC Berkeley from 2009 to 2012. She has been elected to the National Academy of Sciences and Academy of Arts and Sciences. She was President of the Institute of Mathematical Statistics (IMS), Guggenheim Fellow, Tukey Memorial Lecturer of the Bernoulli Society, Rietz Lecturer of IMS, and a COPSS E. L. Scott Prize winner.  She is a member of the UC Berkeley Center for Computational Biology and serves as a scientific advisor at the Simons Institute for the Theory of Computing. She is a Chan-Zuckerberg Biohub Investigator and Weill Neurohub Investigator.

She obtained her BS Degree in Mathematics from Peking University, and MS and PhD Degrees in Statistics from UC Berkeley. She was Assistant Professor at UW-Madison, Visiting Assistant Professor at Yale University, Member of Technical Staff at Lucent Bell-Labs, and Miller Research Professor at Berkeley. She was a Visiting Faculty at MIT, ETH, Poincare Institute, Peking University, INRIA-Paris, Fields Institute at University of Toronto, Newton Institute at Cambridge University, and the Flatiron Institute in NYC.  Recently, she was a 50% consultant researcher in the deep learning group of MSR at Redmond (2022-2023).

Recently, she delivered the Wald Memorial Lectures of IMS and COPSS Distinguished Achievement Award and Lecture (DAAL) (formerly Fisher Award and Lecture) at JSM in 2023. She served on the Inaugural Scientific Committee of the UK Turing Institute for Data Science and AI. She holds an Honorary Doctorate from the University of Lausanne in Switzerland. She is serving on the Editorial Board of Proceedings of National Academy of Sciences (PNAS).

@@

# About Our Research Group

@@

Our research covers several key areas:

## Deep Learning and Machine Learning

Our work in deep learning spans from theoretical understanding to practical applications. Key contributions include mechanistic interpretation through contextual decomposition in transformers, efficient fine-tuning methods like [LoRA+](https://github.com/nikhil-ghosh-berkeley/loraplus), and frameworks for understanding neural network dynamics. We've developed methods like [Adaptive Wavelet Distillation (AWD)](https://github.com/Yu-Group/adaptive-wavelets) for interpreting deep neural networks, and theoretical work on the three stages of dynamics in high-dimensional kernel methods. Our applications range from clinical decision-making in healthcare to cosmological analysis.

Selected recent papers:
- [LoRA+: Efficient Low Rank Adaptation of Large Models](https://arxiv.org/abs/2402.12354) (ICML 2024)
- [Mechanistic Interpretation through Contextual Decomposition in Transformers](https://arxiv.org/pdf/2407.00886) (2024)
- [Diagnosing transformers: illuminating feature space for clinical decision-making](https://arxiv.org/abs/2305.17588) (ICLR 2024)
- [Adaptive Wavelet Distillation from Neural Networks through Interpretation](https://nips.cc/Conferences/2021/ScheduleMultitrack?event=28027) (NeurIPS 2021)

## Veridical Data Science (PCS)

The PCS (Predictability, Computability, Stability) framework represents our approach to responsible, reliable, and transparent data analysis. This framework unifies and expands on best practices in machine learning and statistics, ensuring that data science findings are both scientifically sound and practically actionable. Our work includes applications to genomics, clinical research, and epidemiology, always emphasizing the importance of stability and reproducibility in scientific discovery.

Selected recent papers:
- [Veridical data science](https://www.stat.berkeley.edu/~binyu/ps/papers2020/VDS20-YuKumbier.pdf) (PNAS 2020) - foundational framework paper
- [Epistasis regulates genetic control of cardiac hypertrophy](https://www.medrxiv.org/content/10.1101/2023.11.06.23297858v1) (2023) - PCS application to genomics
- [After Computational Reproducibility: Scientific Reproducibility and Trustworthy AI](https://hdsr.mitpress.mit.edu/pub/8qexde24/release/1) (Harvard Data Science Review 2024)
- [Stable discovery of interpretable subgroups via calibration in causal studies (staDISC)](https://onlinelibrary.wiley.com/doi/10.1111/insr.12427)

## Interpretable Machine Learning

We develop methods to make machine learning models interpretable and trustworthy, particularly for high-stakes applications in healthcare and science. Our work spans contextual decomposition for understanding neural networks, feature importance frameworks, and methods for aligning model explanations with prior knowledge. We emphasize the principle that interpretations should be both faithful to the model and useful for human decision-making.

Selected recent papers:
- [Mechanistic Interpretation through Contextual Decomposition in Transformers](https://arxiv.org/pdf/2407.00886) (2024)
- [MDI+: a flexible random forest-based feature importance framework](https://arxiv.org/abs/2307.01932) (2023)
- [Tell your model where to attend: post-hoc attention steering for LLMs](https://arxiv.org/abs/2311.02262) (ICLR 2024)
- [Definitions, methods, and applications in interpretable machine learning](https://www.stat.berkeley.edu/~binyu/ps/papers2020/iML19-Murdochetal.pdf) (PNAS 2019)

## Tree-based Methods

We develop both theoretical understanding and practical improvements for tree-based methods, including random forests and gradient boosting. Our contributions include the iterative Random Forests (iRF) algorithm for discovering high-order interactions, hierarchical shrinkage methods for improving accuracy and interpretability, and fast interpretable greedy-tree sums (FIGS). We also provide theoretical analysis of when and why tree-based methods succeed or fail.

Selected recent papers:
- [Fast interpretable greedy-tree sums (FIGS)](https://arxiv.org/abs/2201.11931) (2022)
- [Hierarchical shrinkage: improving accuracy and interpretability of tree-based methods](https://arxiv.org/abs/2202.00858) (ICML 2022)
- [Provable Boolean Interaction Recovery from Tree Ensemble obtained via Random Forests](https://arxiv.org/abs/2102.11800) (PNAS 2022)
- [iterative Random Forests to discover predictive and stable high-order interactions](https://www.stat.berkeley.edu/~binyu/ps/papers2018/iRF+SI18.pdf) (PNAS 2018)

<!-- ## Interdisciplinary Research in Biomedicine

Our biomedicine work applies statistical machine learning to problems in genomics, clinical prediction, and biomarker discovery. We collaborate closely with domain experts to develop methods that are both statistically rigorous and clinically meaningful. Recent work includes epistasis analysis for cardiac hypertrophy, metabolomic signatures for pancreatic cancer risk, and clinical AI modeling standards.

Selected recent papers:
- [Epistasis regulates genetic control of cardiac hypertrophy](https://www.medrxiv.org/content/10.1101/2023.11.06.23297858v1) (2023)
- [A blood-based metabolomic signature predictive of risk for pancreatic cancer](https://pubmed.ncbi.nlm.nih.gov/37729870/) (Cell Reports Medicine 2023)
- [Minimum information about clinical artificial intelligence modeling: the MI-CLAIM checklist](https://www.nature.com/articles/s41591-020-1041-y) (Nature Medicine 2020)
- [Unsupervised pattern discovery in spatial gene expression atlas reveals mouse brain regions beyond established ontology](https://www.biorxiv.org/content/10.1101/2023.03.10.531984v2) (2023) -->

@@

My broad vision of data science best described in my article [Veridical Data Science](https://www.pnas.org/content/117/8/3920). In this work, I introduced a framework based on three principles: predictability, computability and stability (abbreviated to PCS). This framework helps guide practitioners who solve domain data problems with data science tools to be creative in their analysis and properly validate their findings. I have written a book on the Veridical Data Science framework together with my former student [Rebecca Barter](https://www.rebeccabarter.com/). The book is being published by the MIT Press in 2024 with a free on-line copy available soon.

### 2025-07-20 index.qmd
Refactor website navigation and update content
Added a new 'People' link to the navigation in _quarto.yml and streamlined the 'Papers' section to a single link. Updated the 'About' page with a comprehensive overview of Professor Bin Yu's background and research focus, while enhancing the teaching section with detailed course descriptions for STAT 154 and STAT 215A. Adjusted styles.css to comment out the background color for the column class, improving layout consistency.

CDSS Chancellor's Distinguished Professor

Departments of [Statistics](https://statistics.berkeley.edu/) and [Electrical Engineering and Computer Sciences](https://eecs.berkeley.edu/), Center for Computational Biology

[UC Berkeley](https://www.berkeley.edu/)

[Chan-Zuckerberg Biohub](https://www.czbiohub.org/) Investigator Alumnus • [Weill Neurohub](https://www.weillneurohub.org/) Investigator

*mail: 367 Evans Hall #3860 • Berkeley, CA 94720*

*phone: 510-642-2781 • fax: 510-642-7892 • [binyu@berkeley.edu](mailto:binyu@berkeley.edu)*

<div class="two-column-container">

<div class="column">
## Recent Publications & Talks

@@

</div>

<div class="column">
## Recent News

@@

</div>

</div>

>>>>

Chancellor's Distinguished Professor of Statistics and Electrical Engineering and Computer Sciences by courtesy

Statistics Department Chair, 2009 - 2012

*mail: 367 Evans Hall #3860 • Berkeley, CA 94720 phone: 510-642-2781 • fax: 510-642-7892 • [binyu@berkeley.edu](mailto:binyu@berkeley.edu)*

## Recent Publications & Talks

@@

## Recent News

@@
